Application of Network Flow and Linear Programming to Scheduling of Clinicians

### 2019-07-18 paper/paper.methods.tex
Clarify adjacency mapping from block to weekend

\subsection{Objectives}
As mentioned briefly in Section \ref{???}, the soft constraints of the clinician scheduling problem are: (1) satisfying clinician block off requests, (2) satisfying clinician weekend off requests, and (3) assigning weekends closer to blocks. We convert these constraints into linear objective functions of the binary variables defined in Section \ref{???}. The first two objectives have a straight-forward formulation as linear functions:
\begin{align}
	&Q_1(X) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{d=1}^{D} (-1)^{\ind(b\,\in\,\mathcal{BR}_c)}\cdot X_{c, b, d} \tag{Block Requests} \label{eqn:obj-block-requests}\\
	&Q_2(Y) = \sum_{c=1}^{C} \sum_{w=1}^{W} (-1)^{\ind(w\,\in\,\mathcal{WR}_c)}\cdot Y_{c, w} \tag{Weekend Requests} \label{eqn:obj-weekend-requests}
\end{align}
where $\ind(\cdot)$ is the indicator function that has value 1 when the condition is met and 0 otherwise. In these two objectives, we penalize any assignments that conflict with a block or weekend request, and aim to maximize the non-conflicting assignments. The Block-Weekend Adjacency is optimized by considering the product $X_{c, b, d}\cdot Y_{c, w}$ for values of $w$ ``adjacent'' to the value of $b$. For instance, if clinician $c$ wants to work during weeks 1 and 2 (that is, block $b = 1$), as well as weekend 1, we would like the product $X_{c, b=1, d} \cdot Y_{c, w=1}$ to be 1, since that indicates both variables are assigned. If at least one of the two variables is not assigned, the product will be 0. This leads to the obvious maximization objective
\begin{align}
	&Q_3(X, Y) = \sum_{min}^{max} ... \tag{Block-Weekend Adjacency} \label{eqn:obj-block-weekend-adj}
\end{align}
However, as it is, $Q_3$ is not a linear function of its variables and cannot be optimized in a linear programming framework. The general approach used to convert such functions into linear objectives is by introducing a helper variable and additional constraints [ref \ref{???}]. In our case, introducing a variable $Z_{c, b, d}$ for every product $X_{c, b, d} \cdot Y_{c, w}$ with $w = ...$, and constraining $Z$ such that 
\begin{align}
	&Z_{c, b, d} \leq X_{c, b, d} \\
	&Z_{c, b, d} \leq Y_{c, w=...} &&\forall d\in\mathcal{D}
\end{align}
allows us to rewrite $Q_3$ as linear function of $Z$,
\begin{align}
	&Q_3(Z) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{d=1}^{D} Z_{c, b, d}
\end{align}
Indeed, whenever $X_{c, b, d} \cdot Y_{c, w} = 1$, $Z_{c, b, d}$ can attain a maximum value of 1, and whenever $X_{c, b, d} \cdot Y_{c, w} = 0$, at least one of $X_{c, b, d}$ or $Y_{c, w}$ must be 0, so $Z_{c, b, d}$ will be constrained to attain a maximum value of 0, giving us the correct adjacency maximization objective. \\

>>>>

\subsection{Objectives}
As mentioned briefly in Section \ref{???}, the soft constraints of the clinician scheduling problem are: (1) satisfying clinician block off requests, (2) satisfying clinician weekend off requests, and (3) assigning weekends closer to blocks. We convert these constraints into linear objective functions of the binary variables defined in Section \ref{???}. The first two objectives have a straight-forward formulation as linear functions:
\begin{align}
	&Q_1(X) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{d=1}^{D} (-1)^{\ind(b\,\in\,\mathcal{BR}_c)}\cdot X_{c, b, d} \tag{Block Requests} \label{eqn:obj-block-requests}\\
	&Q_2(Y) = \sum_{c=1}^{C} \sum_{w=1}^{W} (-1)^{\ind(w\,\in\,\mathcal{WR}_c)}\cdot Y_{c, w} \tag{Weekend Requests} \label{eqn:obj-weekend-requests}
\end{align}
where $\ind(\cdot)$ is the indicator function that has value 1 when the condition is met and 0 otherwise. In these two objectives, we penalize any assignments that conflict with a block or weekend request, and aim to maximize the non-conflicting assignments. The Block-Weekend Adjacency is optimized by considering the product $X_{c, b, d}\cdot Y_{c, w}$ for values of $w$ ``adjacent'' to the value of $b$. For instance, if clinician $c$ wants to work during weeks 1 and 2 (that is, block $b = 1$), as well as weekend 1, we would like the product $X_{c, b=1, d} \cdot Y_{c, w=1}$ to be 1, since that indicates both variables are assigned. If at least one of the two variables is not assigned, the product will be 0. This leads to the obvious maximization objective
\begin{align}
	&Q_3(X, Y) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{d=1}^{D} X_{c, b, d}\cdot Y_{c, w=\varphi(b)} \tag{Block-Weekend Adjacency} \label{eqn:obj-block-weekend-adj}
\end{align}
where $\varphi(b)$ maps a block one-to-one to an adjacent weekend, by whichever definition of adjacency being considered. For instance, in the above example we will have $\varphi(b) = 2b - 1$. \\

However, as it is, $Q_3$ is not a linear function of its variables and cannot be optimized in a linear programming framework. The general approach used to convert such functions into linear objectives is by introducing a helper variable and additional constraints [ref \ref{???}]. In our case, introducing a variable $Z_{c, b, d}$ for every product $X_{c, b, d} \cdot Y_{c, w}$ with $w = \varphi(b)$, and constraining $Z$ such that 
\begin{align}
	&Z_{c, b, d} \leq X_{c, b, d} \\
	&Z_{c, b, d} \leq Y_{c, w=\varphi(b)} &&\forall d\in\mathcal{D}
\end{align}
allows us to rewrite $Q_3$ as linear function of $Z$,
\begin{align}
	&Q_3(Z) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{d=1}^{D} Z_{c, b, d}
\end{align}
Indeed, whenever $X_{c, b, d} \cdot Y_{c, w} = 1$, $Z_{c, b, d}$ can attain a maximum value of 1, and whenever $X_{c, b, d} \cdot Y_{c, w} = 0$, at least one of $X_{c, b, d}$ or $Y_{c, w}$ must be 0, so $Z_{c, b, d}$ will be constrained to attain a maximum value of 0, giving us the correct adjacency maximization objective. \\
